Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-06 Manuscript/manuscript.Rmd
updates

The data sets used for modeling was composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the majority and minority sets. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. For each training set, an $l_{1}$ norm penalized regularization (LASSO) was performed. The $\lambda$ penalization term was tuned over a grid of 10 values on an exponential scale ranging from 0.01 to 10.

@@

We first compared models with either overlap counts or overlap percent predictor types [??? Why not distance?], from the 5kb resolution data (Figure 12). These results were from random forest models fit to data without any re-sampling, using either LASSO or Elastic-Net regulariations. We found that differences in performances across variable reduction techniques were marginal. More importantly, when evaluating model performance using accuracy or AUC, the model appeared to perform relatively well. At second glance, however, when considering both the F1-score and MCC, we were able to assess just how poorly [??? Describe, why is it so?] the model was performing (Supplementary Table 9).

The performances were found to be consistently poor for models with no re-sampling across the other resolutions (Supplementary Figure 13; Supplementary Table 10). The low, and sometimes missing, values indicated that the model was unable to identify many, if any, true positive outcomes. That is, given imbalanced data, the model was unable to predict which genomic bins truly overlapped with a flanked TAD boundary.

>>>>

The data sets used for modeling were composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the majority and minority sets. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. For each training set, an $l_{1}$ norm penalized regularization (LASSO) was performed. The $\lambda$ penalization term was tuned over a grid of 10 values on an exponential scale ranging from 0.01 to 10.

@@

We first assessed performances on random forest models built on fully unbalanced training sets across each resolution, for each predictor type (Figure 5) [??? Why not distance?; !!!included a  more desciptive plot with distance]. When looking at the the conventional metrics used for assessment (accuracy and AUC), models that were not corrected for class imbalance appeared to perform reasonably well.  However, when we considered the F1-score and MCC, we observed a significant degredation in performance across both resolution and predictor type [??? Describe, why is it so?; !!!described]. This was due in part to each models' inability to correctly classify the genomic bins that overlapped with flanked TAD boundaries, as recognized by the small number of true positive cases (Supplementary Table 3). Even without correcting for the class imbalance, we found that, within predictor types, models with distance type predictors consistently outperformed the other two predictor types across all performance metrics.
